---
title: "Markdown"
author: "Me"
date: "31/03/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
data <- read.csv("activity.csv")
head(data)
```

Histogram of the total number of steps taken each day

```{r}
steps_by_day <- aggregate(steps ~ date, data, sum)
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps")
```

Mean and median number of steps taken each day

```{r}
steps_by_day <- aggregate(steps ~ date, data, sum)
rmean <- mean(steps_by_day$steps)
rmedian <- median(steps_by_day$steps)
sprintf("MEAN of steps taken each day = %.3f", rmean)
```


```{r}
sprintf("MEDIAN of steps taken each day = %.3f", rmedian)
```

Time series plot of the average number of steps taken

```{r}
steps_by_interval <- aggregate(steps ~ interval, data, mean)
plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```


The 5-minute interval that, on average, contains the maximum number of steps

```{r}
max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
sprintf("MAXIMUM number of steps in 5-minute interval = %.0f", max_interval)
```

Code to describe and show a strategy for imputing missing data

```{r}
incomplete_data <- sum(!complete.cases(data))
sprintf("MISSING data = %.0f", incomplete_data)
```

The strategy is use the average value for that missing interval. For example, at "date"=="2012-10-01" the value 0 is NA, so the value 1.7169811 will be used as steps. This value came from the data set steps_by_interval.


```{r}
imputed_data <- transform(data, steps = ifelse(is.na(data$steps), steps_by_interval$steps[match(data$interval, steps_by_interval$interval)], data$steps))
```

Histogram of the total number of steps taken each day after missing values are imputed


```{r}
imputed_steps_by_day <- aggregate(steps ~ date, imputed_data, sum)
hist(imputed_steps_by_day$steps, main = paste("Total Steps Each Day"), col="green", xlab="Number of Steps")
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="red", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("green", "red"), lwd=10)
```

MEAN difference


```{r}
rmean <- mean(steps_by_day$steps)
imputed_rmean <- mean(imputed_steps_by_day$steps)
sprintf("MEAN of steps taken each day = %.3f", rmean)
```

```{r}
sprintf("MEAN of steps taken each day with IMPUTED data = %.3f", imputed_rmean)
sprintf("The difference is %.3f ", imputed_rmean-rmean)
```


MEDIAN difference

```{r}
rmedian <- median(steps_by_day$steps)
imputed_rmedian <- median(imputed_steps_by_day$steps)
sprintf("MEDIAN of steps taken each day = %.3f", rmedian)
sprintf("MEDIAN of steps taken each day with IMPUTED data = %.3f", imputed_rmedian)
sprintf("The difference is %.3f ", imputed_rmedian-rmedian)
```
TOTAL steps difference

```{r}
total <- sum(steps_by_day$steps)
imputed_total <- sum(imputed_steps_by_day$steps)
sprintf("TOTAL of steps = %.3f", total)
sprintf("TOTAL of steps with IMPUTED data = %.3f", imputed_total)
sprintf("The difference is %.3f ", imputed_total-total)
```

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
weekend_days_set <- c("Saturday", "Sunday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$date)),weekend_days_set), "Weekend", "Weekday"))
imputed_steps_by_interval <- aggregate(steps ~ interval + dow, imputed_data, mean)
library(lattice)
xyplot(imputed_steps_by_interval$steps ~ imputed_steps_by_interval$interval|imputed_steps_by_interval$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```

